\documentclass[sigconf, authorversion]{acmart}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{algorithmicx}
\usepackage{listings}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{float}
\usepackage{url}
\usepackage{xcolor}
%testtk
\lstdefinestyle{codestyle}{
  basicstyle=\fontsize{9}{11}\selectfont\ttfamily,
  numberstyle=\tiny,
  numbers=left,
  xleftmargin=2em,
  framexleftmargin=1.5em,
  frame=lines,
  breaklines=true,
  tabsize=2,
  escapeinside={<@}{@>},
  captionpos=b,
  language=C++
}
\lstdefinestyle{nolinenumberstyle}{
  basicstyle=\fontsize{9}{11}\selectfont\ttfamily,
  numbers=none,
  xleftmargin=0em,
  framexleftmargin=0em,
  frame=none,
  breaklines=true,
  tabsize=2,
  captionpos=b,
  language=C++
}
\lstset{style=codestyle}
%\renewcommand*{\lstlistoflistings}{List of \lstlistingname s}

%% Commands
\newcommand{\codevar}[1]{\textit{#1}}
\newcommand{\codefn}[1]{\textit{#1}}
\newcommand{\codefile}[1]{\textit{#1}}
\newcommand{\ruleabove}{}
\newcommand{\rulebelow}{}

%Tables 
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\explicitflow}[1]{\vspace{-0.25cm}{\footnotesize\hspace{0.25cm}\textbf{Explicit:} #1 \par}}
\newcommand{\implicitflow}[1]{{\footnotesize\hspace{0.25cm}\textbf{Implicit:} #1 \par}}

\DeclareMathOperator{\pointer}{pointer}
\DeclareMathOperator{\result}{result}
\DeclareMathOperator{\ptrval}{ptrval}
\DeclareMathOperator{\offset}{offset}
\DeclareMathOperator{\myvalue}{value}

% TOG prefers author-name bib system with square brackets
\citestyle{acmnumeric}
\setcitestyle{numbers,sort&compress}


%
% \usepackage[ruled]{algorithm2e} % For algorithms
% \renewcommand{\algorithmcfname}{ALGORITHM}
% \SetAlFnt{\small}
% \SetAlCapFnt{\small}
% \SetAlCapNameFnt{\small}
% \SetAlCapHSkip{0pt}
% \IncMargin{-\parindent}

% Metadata Information
%\acmJournal{TOG}
\acmConference[CCS'19]{Computer and Communications Security}
\acmVolume{26}
\acmYear{2019}
\acmMonth{1}

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{0000001.0000001_2}

% Paper history
\received{February 2019}


% Document starts
\begin{document}
% Title portion
\title{Detecting Non-constant time code in cryptography libraries using a static
  information analysis} 

\author{Adam Mohammed}
\affiliation{%
  \institution{The Pennsylvania State University}
}
\email{aqm5498@psu.edu}

\author{Ernest Defoy III}
\affiliation{%
  \institution{The Pennsylvania State University}
}
\email{edd5103@psu.edu}

\author{Danfeng Zhang}
\affiliation{%
  \institution{The Pennsylvania State University}
}
\email{zhang@psu.edu}

% \renewcommand\shortauthors{Zhou, G. et al}

\begin{abstract}
Identifying timing channels for cryptosystems is often a manual process.
Addressing the cause of the timing channel is only possible once the flaw is
identified. Using information-flow analysis it is possible to identify timing
channels. For example, a branch dependent on secret data may cause a timing
channel if the operations for each path are distinguishable by execution time.
An information flow analysis can track the flow of sensitive information
through a program. The issue is that the number of positives can be very high,
and unequally weighted in the threat that may be posed. A static
information-flow analysis can be used and the precision adjusted to improve
the precision of the analysis.

Improving the precision, means decreasing the overall number of positives while
maintaining all of the higher-risk positives. In this work, an
information flow analysis and a points-to analysis is used in conjunction to
detect secret-dependent branches. Adding field-sensitivity in some cases
improves the precision of the analysis. In some cases, field-sensitivity is not
able to reduce the number of positives, so a white-list can be used to ignore
positives that leak non-sensitive information such as key length.

The baseline analysis combined with the field-sensitivity and whitelist were
used in a case study exploring the effectiveness of each feature on real-world
cryptosystems. The modular exponentiation function was analyzed as it is used
in public-key cryptography and timing channel attacks have been identified in the
code providing this functionality. The results show that improving the precision
can be helpful to reduce the number of positives. An effort to categorize
results is described, to help prioritize results in terms of estimated severity.
\end{abstract}


%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
 \begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002978.10003006.10011608</concept_id>
<concept_desc>Security and privacy~Information flow control</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Security and privacy~Information flow control}
%
% End generated code
%


\keywords{}



\maketitle

\input{introduction}
\input{background}
\input{methodology}
\input{evaluation}
\input{conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{research}
\end{document}
