\documentclass{report}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{algorithmicx}
\usepackage{listings}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{float}
\usepackage{url}
\usepackage{xcolor}
\usepackage{cite}
\usepackage[nottoc]{tocbibind}

\lstdefinestyle{codestyle}{
  basicstyle=\fontsize{9}{11}\selectfont\ttfamily,
  numberstyle=\tiny,
  numbers=left,
  xleftmargin=2em,
  framexleftmargin=1.5em,
  frame=lines,
  breaklines=true,
  tabsize=2,
  escapeinside={<@}{@>},
  captionpos=b,
  language=C++
}
\lstdefinestyle{nolinenumberstyle}{
  basicstyle=\fontsize{9}{11}\selectfont\ttfamily,
  numbers=none,
  xleftmargin=0em,
  framexleftmargin=0em,
  frame=none,
  breaklines=true,
  tabsize=2,
  captionpos=b,
  language=C++
}
\lstset{style=codestyle}
%\renewcommand*{\lstlistoflistings}{List of \lstlistingname s}

%% Commands
\newcommand{\codevar}[1]{\textit{#1}}
\newcommand{\codefn}[1]{\textit{#1}}
\newcommand{\codefile}[1]{\textit{#1}}
\newcommand{\ruleabove}{}
\newcommand{\rulebelow}{}

%Tables 
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\explicitflow}[1]{\vspace{-0.25cm}{\footnotesize\hspace{0.25cm}\textbf{Explicit:} #1 \par}}
\newcommand{\implicitflow}[1]{{\footnotesize\hspace{0.25cm}\textbf{Implicit:} #1 \par}}

\DeclareMathOperator{\pointer}{pointer}
\DeclareMathOperator{\result}{result}
\DeclareMathOperator{\ptrval}{ptrval}
\DeclareMathOperator{\offset}{offset}
\DeclareMathOperator{\myvalue}{value}
\begin{document}
\pagenumbering{roman}
%%%-----------------Title page -------------------
%% Detecting non-constant time code in cryptography libraries using a static
%% information flow analysis
\thispagestyle{empty}
{
  \centering
  {The Pennsylvania State University\par}
  \vspace{0.5cm}
  {The Graduate School\par}
  \vspace{0.5cm}
  {College of Engineering\par}
  \vspace{1.5cm}
  {\uppercase{\textbf{Detecting non-constant time code in}\par}}
  \vspace{1em}
  {\uppercase{\textbf{cryptography libraries using a static}\par}}
  \vspace{1em}
  {\uppercase{\textbf{information flow analysis}\par}}
  \vspace{1.5cm}
  {A Thesis in\par}
  \vspace{0.5cm}
  {Computer Science and Engineering\par}
  \vspace{0.5cm}
  {by\par}
  \vspace{0.5cm}
  {Adam Mohammed\par}
  \vspace{1.5cm}
  {\copyright 2018 Adam Mohammed\par}
  \vspace{1.0cm}
  {Submitted in Partial Fulfillment\par}
  {of the Requirements\par}
  {for the Degree of\par}
  \vspace{1.5cm}
  {Master of Science\par}
  \vspace{1.5cm}
  {December 2018}
  \clearpage
}
%%%------END--------Title page -------------------


%%%-----------------Comittee page -------------------
\noindent
The thesis of Adam Mohammed was reviewed and approved* by the following:

\vspace{1.5cm}
\noindent
Danfeng Zhang\\
Assistant Professor of Computer Science and Engineering\\
Thesis Advisor\\

\vspace{2.5cm}
\noindent
Gang Tan\\
Associate Professor of Computer Science and Engineering\\

\vspace{2.5cm}
\noindent
Chitaranjan Das\\
Distinguished Professor of Computer Science and Engineering\\
Department Head of Computer Science and Engineering\\

\vspace{2.5cm}
\noindent
*Signatures are on file in the Graduate School.\\

\clearpage

%%%------END--------Comittee page -------------------
\begin{center}\large\textbf{Abstract}
  \end{center}

Identifying timing channels for cryptosystems is often a manual process.
Addressing the cause of the timing channel is only possible once the flaw is
identified. Using information-flow analysis it is possible to identify timing
channels. For example, a branch dependent on secret data may cause a timing
channel if the operations for each path are distinguishable by execution time.
An information flow analysis can track the flow of sensitive information
through a program. The issue is that the number of positives can be very high,
and unequally weighted in the threat that may be posed. A static
information-flow analysis can be used and the precision adjusted to improve
the precision of the analysis.

Improving the precision, means decreasing the overall number of positives while
maintaining all of the higher-risk positives. In this work, an
information flow analysis and a points-to analysis is used in conjunction to
detect secret-dependent branches. Adding field-sensitivity in some cases
improves the precision of the analysis. In some cases, field-sensitivity is not
able to reduce the number of positives, so a white-list can be used to ignore
positives that leak non-sensitive information such as key length.

The baseline analysis combined with the field-sensitivity and whitelist were
used in a case study exploring the effectiveness of each feature on real-world
cryptosystems. The modular exponentiation function was analyzed as it is used
in public-key cryptography and timing channel attacks have been identified in the
code providing this functionality. The results show that improving the precision
can be helpful to reduce the number of positives. An effort to categorize
results is described, to help prioritize results in terms of estimated severity.

\clearpage

\tableofcontents
\listoftables
\listoffigures

\clearpage
\chapter*{Acknowledgements}

{This material is based upon work supported by the NSF under Award No. CCF-156641.\par}
\vspace{1em}
\noindent
{Disclaimer: Any opinions, findings, and conclusions or recommendations expressed in this publication are those of the author(s) and do not necessarily reflect the views of the funding agency.\par}

\clearpage
\pagenumbering{arabic}

%%% ------------------ Introduction--------------------
\include{Introduction}
%%% ----END----------- Introduction--------------------

%%% ------------------ Background  --------------------
\include{Background}
%%% ----END----------- Background  --------------------


%%% ------------------ Methodology --------------------
\include{Methodology}
%%% ----END----------- Methodology --------------------

%%% ------------------ Evaluation  --------------------
\include{Evaluation}
%%% ----END----------- Evaluation  --------------------

\include{Conclusion}

\bibliography{research}{}
\bibliographystyle{unsrt}
\end{document}